---
title: "midterm_project"
author: "Mengyuan Li"
date: "10/19/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
library(ggplot2)
library(gridExtra)
```


# problem 1 - Data

```{r}
data = read_excel("data_set.xlsx", skip = 8) 
data = janitor::clean_names(data)
data = mutate(data, sex = factor(sex, levels = c(0,1), 
                                 labels = c("female", "male"))) %>%
                   mutate(eop_size_mm = replace(eop_size_mm, is.na(eop_size_mm), 0),
                          eop_shape = factor(eop_shape))
  


data_1 = data %>%
  group_by(sex) %>%
  ggplot(aes(age, fill = sex)) + geom_histogram() 
data_1


```
Description of Data Cleaning Process: I changed the variables' names to lower-cased and readable names using "janitor clean", and coverted "0,1" under column "sex" to "female" and "male" using function "factor". At the last step, I  replaced all of the missing values by 0 by applying the function "replace". 

Indentification of Key Variables: In the resulting dataset, "sex", "age", and "FHP size" are three key variables. By categoring male and female into different groups, researchers can see how sex relates to exostosis. Also, enthesophyte size might vary by age, which can be checked by detecting ages. Also, "FHP size" is a key because it's the main factor we focus on.

Declare of Distribution: There are `r nrow(data)` participants. By looking at the hitogram I ploted, we can see that the distribution of ages are very similar among males and females. The only difference is that the number of females among all ages is greater than the number of males.

Issues in Existing Dataset: For variable "age_group", it is very confused to represent ages over 60 as "6,7,8". It's enough to use "6" instead of three numbers. Also, the way to covert continuous variables to categorical variables are not correct. For example, 10-15 mm eop size is indicated by 2, and 15-20 mm eop size is indicated by 3. So, for eop size equal to 15, it's a trouble to get corrosponding indicator.



# problem 2 - Visualization

```{r}
data_2 = data %>%
  mutate(age_group = factor(age_group, levels = c(1,2,3,4,5,6,7,8), labels = c("1-17", "18-30", "31-40", "41-50", "51-60", "60+", "60+", "60+"))) %>%
  drop_na() %>%
  filter(!(age_group == "1-17")) %>%
  group_by(age_group, sex) 

p1 = ggplot(data_2, aes(x = fhp_size_mm,  fill = sex)) + geom_density(alpha = .6) + facet_grid(~age_group)
data_2


data_3_1 = data %>%
  mutate(age_group = factor(age_group, levels = c(1,2,3,4,5,6,7,8), labels = c("1-17", "18-30", "31-40", "41-50", "51-60", "60+", "60+", "60+"))) %>%
  filter(!(age_group == "1-17")) %>%
  group_by(age_group, sex) %>%
  summarise(total_people = n()) 

data_3_2 = data %>%
  mutate(age_group = factor(age_group, levels = c(1,2,3,4,5,6,7,8), labels = c("1-17", "18-30", "31-40", "41-50", "51-60", "60+", "60+", "60+"))) %>%
  filter(!(age_group == "1-17")) %>%
  group_by(age_group, sex) %>%
  filter(eop_size != 0) %>%
  summarise(enlarged_eop_people = n()) %>%
  select(enlarged_eop_people)

data_3 = bind_cols(data_3_1, data_3_2) %>%
  mutate(rate = enlarged_eop_people/total_people)

p2 = ggplot(data_3,aes(group = sex, x = age_group, y = rate)) + geom_line(aes(color = sex))
data_3

final_plot = grid.arrange(p1, p2, nrow = 2)
final_plot
 
  
data_4 = data %>%
  mutate(age_group = factor(age_group, levels = c(1,2,3,4,5,6,7,8), labels = c("1-17", "18-30", "31-40", "41-50", "51-60", "60+", "60+", "60+"))) %>%
  filter(!(age_group == "1-17")) %>%
  select(age_group, sex, eop_size_mm, fhp_size_mm) %>% 
  ggplot(aes(x = fhp_size_mm, y = eop_size_mm, color = "pink")) +
  geom_line() +
  facet_grid(vars(sex), vars(age_group)) + 
  viridis::scale_fill_viridis(discrete = TRUE)
data_4
```

Comment: In each group, females have higher average FHP size than males. Males have larger EOP sizes compared to females. Males have higher rate of getting EEOP. I am interested in the question "what is the relationship between EOP size and FPH size for both males and females". 

# problem 3 - Reproducing reported results
```{r}
data_count = data %>%
  mutate(age_group = factor(age_group, levels = c(1,2,3,4,5,6,7,8), labels = c("1-17", "18-30", "31-40", "41-50", "51-60", "60+", "60+", "60+"))) %>%
  group_by(age_group) %>% 
  summarise(number_of_people = n())

data_mean = data %>%
  drop_na() %>%
  filter(age_group != 1) %>%
  group_by(sex) %>%
  select(fhp_size_mm) %>%
  mutate(average_fhp = mean(fhp_size_mm), sd_of_fhp = sd(fhp_size_mm)) %>%
  select(-fhp_size_mm)
data_mean = unique(data_mean, fromLast = TRUE)


data_prevelence = data %>%
  filter(!(eop_size == 0)) %>% 
  filter(! (eop_size == 1)) %>%
  filter(!(age_group == 1))

number_of_diseased_people = nrow(data_prevelence)

total_people = nrow(data)

prevalence = number_of_diseased_people/total_people
prevalence

data_FHP = data %>%
  filter(fhp_size_mm > 40) %>%
  filter(age_group %in% c(6, 7, 8))
number_of_target = nrow(data_FHP)


total_people = data %>%
  filter(age_group %in% c(6,7,8))
total_people = nrow(total_people)

FHP_frequency = number_of_target/total_people
FHP_frequency


  
  
  
```

Comment: the sample size is not consistent with the data. FHP mean(sd) is consistent with my data. EEOP is enlarged EOP. We can use "eop_size" to evaluate this claim. Prevalence is a little bit different from authors' finding. The frequency of FHP is broad trendily consistent with my data. But the specific value is not consistent.  

# problem 4 
There are 2 observations of age 1-17, 303 observations of age 18-30, 204 observations of age 31-40, 207 observations of age 41-50, 200 observations of age 51-60, and 305 observations over 60. The prevalence of EEOP is 57.2%. The frequently of FHP > 40 mm in the over 60s cases is 32.4%. My reports are different from what the authors have. The reason is that my reports are generated by a flawed dataset where many variables are not clearly defined. Based on the evidence, I do not think cell causes horn growth. We need to address the date of specific cell phone using time in order to make strong evidence. 